Structured Versus Decorated Cospans

### 2022-03-08 ResponseToReviews.tex

{\bf We have added a word about Markov processes in Section 6.2, and with this, all the applications of structured or decorated cospans that we know 
of are discussed here.} {\john $\checkmark$} {\chris $\checkmark$}

{\chris Second review starts here}
\item Given the title and abstract of this paper “Structured vs Decorated Cospans” I would expect this paper to provide a judgment that structured  cospans are strictly better than decorated cospans in all circumstances and a justification for that claim, or a detailed analysis of the pros and cons of 
structured vs decorated cospans, some case studies in applications that show how the difference between these double categories is felt in the application domain, and a some description of the tradeoffs.

{\john In fact our goal, as explained in the abstract and introduction, is to prove that the structured and decorated cospan approaches are equivalent under certain conditions, and to illustrate this result with a wide variety of examples.}

{\kenny I think the reader should come to their own conclusion about which framework they should use if they are trying to decide between the two, assuming both are applicable. I wouldn't say that one is better than the other. Personally, I would rather try structured cospans first, because I would rather think about a left adjoint than a symmetric lax monoidal pseudofunctor, but other people might feel differently.}

### 2022-03-08 cospans_new.tex
few more tiny map label changes

Fix a set $\La$ to serve as edge labels.  Define an $\La$-\define{graph} to be a graph $s,t\maps E\to N$ equipped with a function $\ell \maps E \to \La$.  There is a category $\Graph_\La$ where the objects are $\La$-graphs and a morphism from the $\La$-graph 
 \[ \xymatrix{\La & E \ar@<2.5pt>[r]^{s} \ar@<-2.5pt>[r]_{t} \ar[l]_{\ell} & N} \]
 to the $\La$-graph 
\[ \xymatrix{\La & E' \ar@<2.5pt>[r]^{s'} \ar@<-2.5pt>[r]_{t'} \ar[l]_{\ell'} & N'} \]
is a pair of functions $f \maps E \to E', g \maps N \to N'$ such that these diagrams commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$E$};
\node (A') at (0,-1) {$E'$};
\node (B) at (1,0) {$N$};
\node (B') at (1,-1) {$N'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$s$} (B)
(A') edge node[below]{$s'$} (B')
(A) edge node[left]{$f$} (A')
(B) edge node[right]{$g$} (B');

\node (C) at (2,0) {$E$};
\node (C') at (2,-1) {$E'$};
\node (D) at (3,0) {$N$};
\node (D') at (3,-1) {$N'$};
\path[->,font=\scriptsize,>=angle 90]
(C) edge node[above]{$t$} (D)
(C') edge node[below]{$t'$} (D')
(C) edge node[left]{$f$} (C')
(D) edge node[right]{$g$} (D');

\node (E) at (4,-0.5) {$\La$};
\node (G) at (5,0) {$E$};
\node (G') at (5,-1) {$E'$.};
%\node[rotate=-90] at (4.1,-0.5) {$\sim$};
\path[->,font=\scriptsize,>=angle 90]
(G) edge node[above]{$\ell$} (E)
(G) edge node[right]{$f$} (G')
(G') edge node[below]{$\ell'$} (E);
\end{tikzpicture}
\]
There is a functor $U \maps \Graph_\La \to \Fin\Set$ that takes an $\La$-graph to its underlying set of nodes. This has a left adjoint $L \maps \Fin\Set \to \Graph_\La$ sending any set to the $\La$-graph with that set of nodes and no edges.  Both $\Fin\Set$ and $\Graph_\La$ have colimits, and $L$ preserves them.

@@

There is a category $\Petri$ with Petri nets as objects, where a morphism from the Petri net 
$s, t \maps T \to \N[S]$ to the Petri net $s', t' \maps T' \to \N[S']$ is a pair of functions $f \maps T \to T', g \maps S \to S'$ such that the following diagrams commute:
	\[
	\xymatrix{ 
		T \ar[d]_f  \ar[r]^-{s} & \N[S] \ar[d]^-{\N[g]} \\	
		T' \ar[r]_-{s'} & \N[S'] 
	}
	\qquad
	\xymatrix{ 
		T \ar[d]_f  \ar[r]^-{t} & \N[S] \ar[d]^-{\N[g]} \\	
		T' \ar[r]_-{t'} & \N[S'] . 
	}
	\]
There is a functor $R \maps \Petri \to \Fin\Set$ sending any Petri net to its set of places, and this has a left adjoint $L \maps \Fin\Set \to \Petri$ sending any finite set $S$ to the Petri net with $S$ as its set of places and no transitions \cite[Lemma 11]{BM}.   Since both $\Fin\Set$ and $\Petri$ have finite colimits and $L$ preserves them, \cref{thm:structured_cospans} yields a symmetric monoidal double category ${}_L \lCsp(\Petri)$ in which:
\begin{itemize}
\item objects are finite sets,
\item vertical 1-morphisms are functions,
\item horizontal 1-cells are \define{open Petri nets}, which are cospans in $\Petri$ of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (D) at (-3,0) {$L(X)$};
\node (E) at (-2,0) {$P$};
\node (F) at (-1,0) {$L(Y)$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above] {$i$} (E)
(F) edge node [above] {$o$} (E);
\end{tikzpicture}
\]
\item 2-morphisms are diagrams in $\Petri$ of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (E) at (3,0) {$L(X)$};
\node (F) at (5,0) {$L(Y)$};
\node (G) at (4,0) {$P$};
\node (E') at (3,-1) {$L(X')$};
\node (F') at (5,-1) {$L(Y')$.};
\node (G') at (4,-1) {$P'$};
\path[->,font=\scriptsize,>=angle 90]
(F) edge node[above]{$o$} (G)
(E) edge node[left]{$L(f)$} (E')
(F) edge node[right]{$L(g)$} (F')
(G) edge node[left]{$\alpha$} (G')
(E) edge node[above]{$i$} (G)
(E') edge node[below]{$i'$} (G')
(F') edge node[below]{$o'$} (G');
\end{tikzpicture}
\]
\end{itemize}

@@

Now we turn to the details.  A \define{Petri net with rates} is a Petri net $s,t \maps T \to \N[S]$ together with a function $r \maps T \to (0,\infty)$ assigning to each transition $\tau \in T$ a positive real number called its \define{rate constant}.  There is a category $\Petri_r$ whose objects are Petri nets with rates, where a morphism from 
\[   \xymatrix{ (0,\infty) & T \ar[l]_-r \ar@<-.5ex>[r]_-t \ar@<.5ex>[r]^-s & \N[S] }\]
 to 
 \[   \xymatrix{ (0,\infty) & T' \ar[l]_-{r'} \ar@<-.5ex>[r]_-{t'} \ar@<.5ex>[r]^-{s'} & \N[S'] }\]
is a morphism of the underlying Petri nets such that the following diagram also commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (E) at (3,-0.5) {$(0,\infty)$};
\node (G) at (4,0) {$T$};
\node (G') at (4,-1) {$T'$};
\path[->,font=\scriptsize,>=angle 90]
(G) edge node[right]{$f$} (G')
(G) edge node[above]{$r$} (E)
(G') edge node[below]{$r'$} (E);
\end{tikzpicture}
\]
There is a functor $R \maps \Petri_r \to \Set$ sending any Petri net with rates to its set of places, and this has a left adjoint $L \maps \Set \to \Petri_r$ sending any set $S$ to the Petri net with $S$ as its set of places and no transitions \cite[Lemma 6.18]{BC}.   Since $\Petri_r$ has finite colimits \cite[Lemma 6.19]{BC}, it follows that there is a symmetric monoidal double category ${}_L \lCsp(\Petri_r)$ where:
\begin{itemize}
\item objects are finite sets,
\item vertical 1-morphisms are functions,
\item horizontal 1-cells are \define{open Petri nets with rates}, namely diagrams in $\Petri_r$ of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (D) at (-3,0) {$L(X)$};
\node (E) at (-2,0) {$P$};
\node (F) at (-1,0) {$L(Y)$,};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above] {$i$} (E)
(F) edge node [above] {$o$} (E);
\end{tikzpicture}
\]
\item 2-morphisms are diagrams in $\Petri_r$ of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (E) at (3,0) {$L(X)$};
\node (F) at (5,0) {$L(Y)$};
\node (G) at (4,0) {$P$};
\node (E') at (3,-1) {$L(X')$};
\node (F') at (5,-1) {$L(Y')$.};
\node (G') at (4,-1) {$P'$};
\path[->,font=\scriptsize,>=angle 90]
(F) edge node[above]{$i$} (G)
(E) edge node[left]{$L(f)$} (E')
(F) edge node[right]{$L(g)$} (F')
(G) edge node[left]{$\alpha$} (G')
(E) edge node[above]{$o$} (G)
(E') edge node[below]{$i'$} (G')
(F') edge node[below]{$o'$} (G');
\end{tikzpicture}
\]
\end{itemize}

>>>>

Fix a set $\La$ to serve as edge labels.  Define an $\La$-\define{graph} to be a graph $s,t\maps E\to N$ equipped with a function $\ell \maps E \to \La$.  There is a category $\Graph_\La$ where the objects are $\La$-graphs and a morphism from the $\La$-graph 
 \[ \xymatrix{\La & E \ar@<2.5pt>[r]^{s} \ar@<-2.5pt>[r]_{t} \ar[l]_{\ell} & N} \]
 to the $\La$-graph 
\[ \xymatrix{\La & E' \ar@<2.5pt>[r]^{s'} \ar@<-2.5pt>[r]_{t'} \ar[l]_{\ell'} & N'} \]
is a pair of functions $f \maps N \to N', g \maps E \to E'$ such that these diagrams commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$E$};
\node (A') at (0,-1) {$E'$};
\node (B) at (1,0) {$N$};
\node (B') at (1,-1) {$N'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$s$} (B)
(A') edge node[below]{$s'$} (B')
(A) edge node[left]{$g$} (A')
(B) edge node[right]{$f$} (B');

\node (C) at (2,0) {$E$};
\node (C') at (2,-1) {$E'$};
\node (D) at (3,0) {$N$};
\node (D') at (3,-1) {$N'$};
\path[->,font=\scriptsize,>=angle 90]
(C) edge node[above]{$t$} (D)
(C') edge node[below]{$t'$} (D')
(C) edge node[left]{$g$} (C')
(D) edge node[right]{$f$} (D');

\node (E) at (4,-0.5) {$\La$};
\node (G) at (5,0) {$E$};
\node (G') at (5,-1) {$E'$.};
%\node[rotate=-90] at (4.1,-0.5) {$\sim$};
\path[->,font=\scriptsize,>=angle 90]
(G) edge node[above]{$\ell$} (E)
(G) edge node[right]{$g$} (G')
(G') edge node[below]{$\ell'$} (E);
\end{tikzpicture}
\]
There is a functor $U \maps \Graph_\La \to \Fin\Set$ that takes an $\La$-graph to its underlying set of nodes. This has a left adjoint $L \maps \Fin\Set \to \Graph_\La$ sending any set to the $\La$-graph with that set of nodes and no edges.  Both $\Fin\Set$ and $\Graph_\La$ have colimits, and $L$ preserves them.

@@

There is a category $\Petri$ with Petri nets as objects, where a morphism from the Petri net 
$s, t \maps T \to \N[S]$ to the Petri net $s', t' \maps T' \to \N[S']$ is a pair of functions $f \maps S \to S', g \maps T \to T'$ such that the following diagrams commute:
	\[
	\xymatrix{ 
		T \ar[d]_g  \ar[r]^-{s} & \N[S] \ar[d]^-{\N[f]} \\	
		T' \ar[r]_-{s'} & \N[S'] 
	}
	\qquad
	\xymatrix{ 
		T \ar[d]_g  \ar[r]^-{t} & \N[S] \ar[d]^-{\N[f]} \\	
		T' \ar[r]_-{t'} & \N[S'] . 
	}
	\]
There is a functor $R \maps \Petri \to \Fin\Set$ sending any Petri net to its set of places, and this has a left adjoint $L \maps \Fin\Set \to \Petri$ sending any finite set $S$ to the Petri net with $S$ as its set of places and no transitions \cite[Lemma 11]{BM}.   Since both $\Fin\Set$ and $\Petri$ have finite colimits and $L$ preserves them, \cref{thm:structured_cospans} yields a symmetric monoidal double category ${}_L \lCsp(\Petri)$ in which:
\begin{itemize}
\item objects are finite sets,
\item vertical 1-morphisms are functions,
\item horizontal 1-cells are \define{open Petri nets}, which are cospans in $\Petri$ of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (D) at (-3,0) {$L(X)$};
\node (E) at (-2,0) {$P$};
\node (F) at (-1,0) {$L(Y)$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above] {$i$} (E)
(F) edge node [above] {$o$} (E);
\end{tikzpicture}
\]
\item 2-morphisms are diagrams in $\Petri$ of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (E) at (3,0) {$L(X)$};
\node (F) at (5,0) {$L(Y)$};
\node (G) at (4,0) {$P$};
\node (E') at (3,-1) {$L(X')$};
\node (F') at (5,-1) {$L(Y')$.};
\node (G') at (4,-1) {$P'$};
\path[->,font=\scriptsize,>=angle 90]
(F) edge node[above]{$o$} (G)
(E) edge node[left]{$L(f)$} (E')
(F) edge node[right]{$L(g)$} (F')
(G) edge node[left]{$\alpha$} (G')
(E) edge node[above]{$i$} (G)
(E') edge node[below]{$i'$} (G')
(F') edge node[below]{$o'$} (G');
\end{tikzpicture}
\]
\end{itemize}

@@

Now we turn to the details.  A \define{Petri net with rates} is a Petri net $s,t \maps T \to \N[S]$ together with a function $r \maps T \to (0,\infty)$ assigning to each transition $\tau \in T$ a positive real number called its \define{rate constant}.  There is a category $\Petri_r$ whose objects are Petri nets with rates, where a morphism from 
\[   \xymatrix{ (0,\infty) & T \ar[l]_-r \ar@<-.5ex>[r]_-t \ar@<.5ex>[r]^-s & \N[S] }\]
 to 
 \[   \xymatrix{ (0,\infty) & T' \ar[l]_-{r'} \ar@<-.5ex>[r]_-{t'} \ar@<.5ex>[r]^-{s'} & \N[S'] }\]
is a morphism of the underlying Petri nets such that the following diagram also commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (E) at (3,-0.5) {$(0,\infty)$};
\node (G) at (4,0) {$T$};
\node (G') at (4,-1) {$T'$};
\path[->,font=\scriptsize,>=angle 90]
(G) edge node[right]{$g$} (G')
(G) edge node[above]{$r$} (E)
(G') edge node[below]{$r'$} (E);
\end{tikzpicture}
\]
There is a functor $R \maps \Petri_r \to \Set$ sending any Petri net with rates to its set of places, and this has a left adjoint $L \maps \Set \to \Petri_r$ sending any set $S$ to the Petri net with $S$ as its set of places and no transitions \cite[Lemma 6.18]{BC}.   Since $\Petri_r$ has finite colimits \cite[Lemma 6.19]{BC}, it follows that there is a symmetric monoidal double category ${}_L \lCsp(\Petri_r)$ where:
\begin{itemize}
\item objects are finite sets,
\item vertical 1-morphisms are functions,
\item horizontal 1-cells are \define{open Petri nets with rates}, namely diagrams in $\Petri_r$ of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (D) at (-3,0) {$L(X)$};
\node (E) at (-2,0) {$P$};
\node (F) at (-1,0) {$L(Y)$,};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above] {$i$} (E)
(F) edge node [above] {$o$} (E);
\end{tikzpicture}
\]
\item 2-morphisms are diagrams in $\Petri_r$ of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (E) at (3,0) {$L(X)$};
\node (F) at (5,0) {$L(Y)$};
\node (G) at (4,0) {$P$};
\node (E') at (3,-1) {$L(X')$};
\node (F') at (5,-1) {$L(Y')$.};
\node (G') at (4,-1) {$P'$};
\path[->,font=\scriptsize,>=angle 90]
(F) edge node[above]{$i$} (G)
(E) edge node[left]{$L(f)$} (E')
(F) edge node[right]{$L(g)$} (F')
(G) edge node[left]{$\alpha$} (G')
(E) edge node[above]{$o$} (G)
(E') edge node[below]{$i'$} (G')
(F') edge node[below]{$o'$} (G');
\end{tikzpicture}
\]
\end{itemize}
